### 2024-03-16 img/plot_thinning.tex
update

% Created by tikzDevice version 0.12.3.1 on 2022-07-05 16:45:06
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,216.81);
\begin{scope}
\path[clip] ( 20.72, 20.72) rectangle (355.85,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 20.72, 40.21) --
	(355.85, 40.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 20.72, 74.05) --
	(355.85, 74.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 20.72,107.89) --
	(355.85,107.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 20.72,141.73) --
	(355.85,141.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 20.72,175.57) --
	(355.85,175.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.04, 20.72) --
	( 44.04,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.44, 20.72) --
	(111.44,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.85, 20.72) --
	(178.85,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (246.25, 20.72) --
	(246.25,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.65, 20.72) --
	(313.65,211.31);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- (340.62, 40.21);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] (335.69, 37.36) --
	(340.62, 40.21) --
	(335.69, 43.05) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 44.04, 40.21) -- ( 44.04,202.65);

\path[draw=drawColor,line width= 0.3pt,line join=round,fill=fillColor] ( 46.88,197.72) --
	( 44.04,202.65) --
	( 41.19,197.72) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.04,141.73) --
	( 45.39,147.15) --
	( 46.73,149.38) --
	( 48.08,151.09) --
	( 49.43,152.53) --
	( 50.78,153.78) --
	( 52.13,154.90) --
	( 53.47,155.92) --
	( 54.82,156.85) --
	( 56.17,157.71) --
	( 57.52,158.51) --
	( 58.87,159.26) --
	( 60.22,159.95) --
	( 61.56,160.60) --
	( 62.91,161.20) --
	( 64.26,161.77) --
	( 65.61,162.29) --
	( 66.96,162.78) --
	( 68.30,163.23) --
	( 69.65,163.65) --
	( 71.00,164.04) --
	( 72.35,164.39) --
	( 73.70,164.71) --
	( 75.04,165.00) --
	( 76.39,165.26) --
	( 77.74,165.49) --
	( 79.09,165.69) --
	( 80.44,165.86) --
	( 81.78,166.01) --
	( 83.13,166.12) --
	( 84.48,166.21) --
	( 85.83,166.27) --
	( 87.18,166.30) --
	( 88.53,166.30) --
	( 89.87,166.28) --
	( 91.22,166.23) --
	( 92.57,166.16) --
	( 93.92,166.05) --
	( 95.27,165.93) --
	( 96.61,165.77) --
	( 97.96,165.59) --
	( 99.31,165.39) --
	(100.66,165.15) --
	(102.01,164.90) --
	(103.35,164.62) --
	(104.70,164.31) --
	(106.05,163.98) --
	(107.40,163.63) --
	(108.75,163.25) --
	(110.09,162.84) --
	(111.44,162.42) --
	(112.79,161.97) --
	(114.14,161.50) --
	(115.49,161.00) --
	(116.83,160.49) --
	(118.18,159.95) --
	(119.53,159.39) --
	(120.88,158.80) --
	(122.23,158.20) --
	(123.58,157.58) --
	(124.92,156.93) --
	(126.27,156.27) --
	(127.62,155.58) --
	(128.97,154.88) --
	(130.32,154.15) --
	(131.66,153.41) --
	(133.01,152.65) --
	(134.36,151.87) --
	(135.71,151.08) --
	(137.06,150.26) --
	(138.40,149.43) --
	(139.75,148.59) --
	(141.10,147.72) --
	(142.45,146.85) --
	(143.80,145.96) --
	(145.14,145.05) --
	(146.49,144.13) --
	(147.84,143.20) --
	(149.19,142.25) --
	(150.54,141.29) --
	(151.89,140.32) --
	(153.23,139.34) --
	(154.58,138.34) --
	(155.93,137.34) --
	(157.28,136.32) --
	(158.63,135.30) --
	(159.97,134.27) --
	(161.32,133.23) --
	(162.67,132.18) --
	(164.02,131.12) --
	(165.37,130.06) --
	(166.71,128.99) --
	(168.06,127.92) --
	(169.41,126.84) --
	(170.76,125.76) --
	(172.11,124.67) --
	(173.45,123.58) --
	(174.80,122.49) --
	(176.15,121.39) --
	(177.50,120.30) --
	(178.85,119.20) --
	(180.19,118.10) --
	(181.54,117.01) --
	(182.89,115.91) --
	(184.24,114.81) --
	(185.59,113.72) --
	(186.94,112.63) --
	(188.28,111.55) --
	(189.63,110.46) --
	(190.98,109.39) --
	(192.33,108.31) --
	(193.68,107.25) --
	(195.02,106.18) --
	(196.37,105.13) --
	(197.72,104.08) --
	(199.07,103.05) --
	(200.42,102.02) --
	(201.76,100.99) --
	(203.11, 99.98) --
	(204.46, 98.98) --
	(205.81, 97.99) --
	(207.16, 97.02) --
	(208.50, 96.05) --
	(209.85, 95.10) --
	(211.20, 94.16) --
	(212.55, 93.23) --
	(213.90, 92.32) --
	(215.24, 91.43) --
	(216.59, 90.55) --
	(217.94, 89.68) --
	(219.29, 88.83) --
	(220.64, 88.00) --
	(221.99, 87.19) --
	(223.33, 86.39) --
	(224.68, 85.61) --
	(226.03, 84.86) --
	(227.38, 84.12) --
	(228.73, 83.40) --
	(230.07, 82.70) --
	(231.42, 82.02) --
	(232.77, 81.37) --
	(234.12, 80.73) --
	(235.47, 80.12) --
	(236.81, 79.53) --
	(238.16, 78.97) --
	(239.51, 78.43) --
	(240.86, 77.91) --
	(242.21, 77.41) --
	(243.55, 76.94) --
	(244.90, 76.50) --
	(246.25, 76.08) --
	(247.60, 75.69) --
	(248.95, 75.32) --
	(250.30, 74.98) --
	(251.64, 74.67) --
	(252.99, 74.38) --
	(254.34, 74.12) --
	(255.69, 73.89) --
	(257.04, 73.68) --
	(258.38, 73.51) --
	(259.73, 73.36) --
	(261.08, 73.24) --
	(262.43, 73.15) --
	(263.78, 73.09) --
	(265.12, 73.06) --
	(266.47, 73.05) --
	(267.82, 73.08) --
	(269.17, 73.13) --
	(270.52, 73.22) --
	(271.86, 73.33) --
	(273.21, 73.48) --
	(274.56, 73.65) --
	(275.91, 73.86) --
	(277.26, 74.09) --
	(278.60, 74.36) --
	(279.95, 74.66) --
	(281.30, 74.98) --
	(282.65, 75.34) --
	(284.00, 75.72) --
	(285.35, 76.14) --
	(286.69, 76.59) --
	(288.04, 77.07) --
	(289.39, 77.57) --
	(290.74, 78.11) --
	(292.09, 78.68) --
	(293.43, 79.27) --
	(294.78, 79.90) --
	(296.13, 80.56) --
	(297.48, 81.24) --
	(298.83, 81.96) --
	(300.17, 82.70) --
	(301.52, 83.47) --
	(302.87, 84.27) --
	(304.22, 85.10) --
	(305.57, 85.96) --
	(306.91, 86.84) --
	(308.26, 87.76) --
	(309.61, 88.70) --
	(310.96, 89.67) --
	(312.31, 90.66) --
	(313.65, 91.68);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.04,175.57) --
	( 45.39,175.57) --
	( 46.73,175.57) --
	( 48.08,175.57) --
	( 49.43,175.57) --
	( 50.78,175.57) --
	( 52.13,175.57) --
	( 53.47,175.57) --
	( 54.82,175.57) --
	( 56.17,175.57) --
	( 57.52,175.57) --
	( 58.87,175.57) --
	( 60.22,175.57) --
	( 61.56,175.57) --
	( 62.91,175.57) --
	( 64.26,175.57) --
	( 65.61,175.57) --
	( 66.96,175.57) --
	( 68.30,175.57) --
	( 69.65,175.57) --
	( 71.00,175.57) --
	( 72.35,175.57) --
	( 73.70,175.57) --
	( 75.04,175.57) --
	( 76.39,175.57) --
	( 77.74,175.57) --
	( 79.09,175.57) --
	( 80.44,175.57) --
	( 81.78,175.57) --
	( 83.13,175.57) --
	( 84.48,175.57) --
	( 85.83,175.57) --
	( 87.18,175.57) --
	( 88.53,175.57) --
	( 89.87,175.57) --
	( 91.22,175.57) --
	( 92.57,175.57) --
	( 93.92,175.57) --
	( 95.27,175.57) --
	( 96.61,175.57) --
	( 97.96,175.57) --
	( 99.31,175.57) --
	(100.66,175.57) --
	(102.01,175.57) --
	(103.35,175.57) --
	(104.70,175.57) --
	(106.05,175.57) --
	(107.40,175.57) --
	(108.75,175.57) --
	(110.09,175.57) --
	(111.44,175.57) --
	(112.79,175.57) --
	(114.14,175.57) --
	(115.49,175.57) --
	(116.83,175.57) --
	(118.18,175.57) --
	(119.53,175.57) --
	(120.88,175.57) --
	(122.23,175.57) --
	(123.58,175.57) --
	(124.92,175.57) --
	(126.27,175.57) --
	(127.62,175.57) --
	(128.97,175.57) --
	(130.32,175.57) --
	(131.66,175.57) --
	(133.01,175.57) --
	(134.36,175.57) --
	(135.71,175.57) --
	(137.06,175.57) --
	(138.40,175.57) --
	(139.75,175.57) --
	(141.10,175.57) --
	(142.45,175.57) --
	(143.80,175.57) --
	(145.14,175.57) --
	(146.49,175.57) --
	(147.84,175.57) --
	(149.19,175.57) --
	(150.54,175.57) --
	(151.89,175.57) --
	(153.23,175.57) --
	(154.58,175.57) --
	(155.93,175.57) --
	(157.28,175.57) --
	(158.63,175.57) --
	(159.97,175.57) --
	(161.32,175.57) --
	(162.67,175.57) --
	(164.02,175.57) --
	(165.37,175.57) --
	(166.71,175.57) --
	(168.06,175.57) --
	(169.41,175.57) --
	(170.76,175.57) --
	(172.11,175.57) --
	(173.45,175.57) --
	(174.80,175.57) --
	(176.15,175.57) --
	(177.50,175.57) --
	(178.85,175.57) --
	(180.19,175.57) --
	(181.54,175.57) --
	(182.89,175.57) --
	(184.24,175.57) --
	(185.59,175.57) --
	(186.94,175.57) --
	(188.28,175.57) --
	(189.63,175.57) --
	(190.98,175.57) --
	(192.33,175.57) --
	(193.68,175.57) --
	(195.02,175.57) --
	(196.37,175.57) --
	(197.72,175.57) --
	(199.07,175.57) --
	(200.42,175.57) --
	(201.76,175.57) --
	(203.11,175.57) --
	(204.46,175.57) --
	(205.81,175.57) --
	(207.16,175.57) --
	(208.50,175.57) --
	(209.85,175.57) --
	(211.20,175.57) --
	(212.55,175.57) --
	(213.90,175.57) --
	(215.24,175.57) --
	(216.59,175.57) --
	(217.94,175.57) --
	(219.29,175.57) --
	(220.64,175.57) --
	(221.99,175.57) --
	(223.33,175.57) --
	(224.68,175.57) --
	(226.03,175.57) --
	(227.38,175.57) --
	(228.73,175.57) --
	(230.07,175.57) --
	(231.42,175.57) --
	(232.77,175.57) --
	(234.12,175.57) --
	(235.47,175.57) --
	(236.81,175.57) --
	(238.16,175.57) --
	(239.51,175.57) --
	(240.86,175.57) --
	(242.21,175.57) --
	(243.55,175.57) --
	(244.90,175.57) --
	(246.25,175.57) --
	(247.60,175.57) --
	(248.95,175.57) --
	(250.30,175.57) --
	(251.64,175.57) --
	(252.99,175.57) --
	(254.34,175.57) --
	(255.69,175.57) --
	(257.04,175.57) --
	(258.38,175.57) --
	(259.73,175.57) --
	(261.08,175.57) --
	(262.43,175.57) --
	(263.78,175.57) --
	(265.12,175.57) --
	(266.47,175.57) --
	(267.82,175.57) --
	(269.17,175.57) --
	(270.52,175.57) --
	(271.86,175.57) --
	(273.21,175.57) --
	(274.56,175.57) --
	(275.91,175.57) --
	(277.26,175.57) --
	(278.60,175.57) --
	(279.95,175.57) --
	(281.30,175.57) --
	(282.65,175.57) --
	(284.00,175.57) --
	(285.35,175.57) --
	(286.69,175.57) --
	(288.04,175.57) --
	(289.39,175.57) --
	(290.74,175.57) --
	(292.09,175.57) --
	(293.43,175.57) --
	(294.78,175.57) --
	(296.13,175.57) --
	(297.48,175.57) --
	(298.83,175.57) --
	(300.17,175.57) --
	(301.52,175.57) --
	(302.87,175.57) --
	(304.22,175.57) --
	(305.57,175.57) --
	(306.91,175.57) --
	(308.26,175.57) --
	(309.61,175.57) --
	(310.96,175.57) --
	(312.31,175.57) --
	(313.65,175.57);
\definecolor{fillColor}{RGB}{255,0,0}

\path[fill=fillColor,fill opacity=0.10] ( 44.04,175.57) --
	( 45.39,175.57) --
	( 46.73,175.57) --
	( 48.08,175.57) --
	( 49.43,175.57) --
	( 50.78,175.57) --
	( 52.13,175.57) --
	( 53.47,175.57) --
	( 54.82,175.57) --
	( 56.17,175.57) --
	( 57.52,175.57) --
	( 58.87,175.57) --
	( 60.22,175.57) --
	( 61.56,175.57) --
	( 62.91,175.57) --
	( 64.26,175.57) --
	( 65.61,175.57) --
	( 66.96,175.57) --
	( 68.30,175.57) --
	( 69.65,175.57) --
	( 71.00,175.57) --
	( 72.35,175.57) --
	( 73.70,175.57) --
	( 75.04,175.57) --
	( 76.39,175.57) --
	( 77.74,175.57) --
	( 79.09,175.57) --
	( 80.44,175.57) --
	( 81.78,175.57) --
	( 83.13,175.57) --
	( 84.48,175.57) --
	( 85.83,175.57) --
	( 87.18,175.57) --
	( 88.53,175.57) --
	( 89.87,175.57) --
	( 91.22,175.57) --
	( 92.57,175.57) --
	( 93.92,175.57) --
	( 95.27,175.57) --
	( 96.61,175.57) --
	( 97.96,175.57) --
	( 99.31,175.57) --
	(100.66,175.57) --
	(102.01,175.57) --
	(103.35,175.57) --
	(104.70,175.57) --
	(106.05,175.57) --
	(107.40,175.57) --
	(108.75,175.57) --
	(110.09,175.57) --
	(111.44,175.57) --
	(112.79,175.57) --
	(114.14,175.57) --
	(115.49,175.57) --
	(116.83,175.57) --
	(118.18,175.57) --
	(119.53,175.57) --
	(120.88,175.57) --
	(122.23,175.57) --
	(123.58,175.57) --
	(124.92,175.57) --
	(126.27,175.57) --
	(127.62,175.57) --
	(128.97,175.57) --
	(130.32,175.57) --
	(131.66,175.57) --
	(133.01,175.57) --
	(134.36,175.57) --
	(135.71,175.57) --
	(137.06,175.57) --
	(138.40,175.57) --
	(139.75,175.57) --
	(141.10,175.57) --
	(142.45,175.57) --
	(143.80,175.57) --
	(145.14,175.57) --
	(146.49,175.57) --
	(147.84,175.57) --
	(149.19,175.57) --
	(150.54,175.57) --
	(151.89,175.57) --
	(153.23,175.57) --
	(154.58,175.57) --
	(155.93,175.57) --
	(157.28,175.57) --
	(158.63,175.57) --
	(159.97,175.57) --
	(161.32,175.57) --
	(162.67,175.57) --
	(164.02,175.57) --
	(165.37,175.57) --
	(166.71,175.57) --
	(168.06,175.57) --
	(169.41,175.57) --
	(170.76,175.57) --
	(172.11,175.57) --
	(173.45,175.57) --
	(174.80,175.57) --
	(176.15,175.57) --
	(177.50,175.57) --
	(178.85,175.57) --
	(180.19,175.57) --
	(181.54,175.57) --
	(182.89,175.57) --
	(184.24,175.57) --
	(185.59,175.57) --
	(186.94,175.57) --
	(188.28,175.57) --
	(189.63,175.57) --
	(190.98,175.57) --
	(192.33,175.57) --
	(193.68,175.57) --
	(195.02,175.57) --
	(196.37,175.57) --
	(197.72,175.57) --
	(199.07,175.57) --
	(200.42,175.57) --
	(201.76,175.57) --
	(203.11,175.57) --
	(204.46,175.57) --
	(205.81,175.57) --
	(207.16,175.57) --
	(208.50,175.57) --
	(209.85,175.57) --
	(211.20,175.57) --
	(212.55,175.57) --
	(213.90,175.57) --
	(215.24,175.57) --
	(216.59,175.57) --
	(217.94,175.57) --
	(219.29,175.57) --
	(220.64,175.57) --
	(221.99,175.57) --
	(223.33,175.57) --
	(224.68,175.57) --
	(226.03,175.57) --
	(227.38,175.57) --
	(228.73,175.57) --
	(230.07,175.57) --
	(231.42,175.57) --
	(232.77,175.57) --
	(234.12,175.57) --
	(235.47,175.57) --
	(236.81,175.57) --
	(238.16,175.57) --
	(239.51,175.57) --
	(240.86,175.57) --
	(242.21,175.57) --
	(243.55,175.57) --
	(244.90,175.57) --
	(246.25,175.57) --
	(247.60,175.57) --
	(248.95,175.57) --
	(250.30,175.57) --
	(251.64,175.57) --
	(252.99,175.57) --
	(254.34,175.57) --
	(255.69,175.57) --
	(257.04,175.57) --
	(258.38,175.57) --
	(259.73,175.57) --
	(261.08,175.57) --
	(262.43,175.57) --
	(263.78,175.57) --
	(265.12,175.57) --
	(266.47,175.57) --
	(267.82,175.57) --
	(269.17,175.57) --
	(270.52,175.57) --
	(271.86,175.57) --
	(273.21,175.57) --
	(274.56,175.57) --
	(275.91,175.57) --
	(277.26,175.57) --
	(278.60,175.57) --
	(279.95,175.57) --
	(281.30,175.57) --
	(282.65,175.57) --
	(284.00,175.57) --
	(285.35,175.57) --
	(286.69,175.57) --
	(288.04,175.57) --
	(289.39,175.57) --
	(290.74,175.57) --
	(292.09,175.57) --
	(293.43,175.57) --
	(294.78,175.57) --
	(296.13,175.57) --
	(297.48,175.57) --
	(298.83,175.57) --
	(300.17,175.57) --
	(301.52,175.57) --
	(302.87,175.57) --
	(304.22,175.57) --
	(305.57,175.57) --
	(306.91,175.57) --
	(308.26,175.57) --
	(309.61,175.57) --
	(310.96,175.57) --
	(312.31,175.57) --
	(313.65,175.57) --
	(313.65, 91.68) --
	(312.31, 90.66) --
	(310.96, 89.67) --
	(309.61, 88.70) --
	(308.26, 87.76) --
	(306.91, 86.84) --
	(305.57, 85.96) --
	(304.22, 85.10) --
	(302.87, 84.27) --
	(301.52, 83.47) --
	(300.17, 82.70) --
	(298.83, 81.96) --
	(297.48, 81.24) --
	(296.13, 80.56) --
	(294.78, 79.90) --
	(293.43, 79.27) --
	(292.09, 78.68) --
	(290.74, 78.11) --
	(289.39, 77.57) --
	(288.04, 77.07) --
	(286.69, 76.59) --
	(285.35, 76.14) --
	(284.00, 75.72) --
	(282.65, 75.34) --
	(281.30, 74.98) --
	(279.95, 74.66) --
	(278.60, 74.36) --
	(277.26, 74.09) --
	(275.91, 73.86) --
	(274.56, 73.65) --
	(273.21, 73.48) --
	(271.86, 73.33) --
	(270.52, 73.22) --
	(269.17, 73.13) --
	(267.82, 73.08) --
	(266.47, 73.05) --
	(265.12, 73.06) --
	(263.78, 73.09) --
	(262.43, 73.15) --
	(261.08, 73.24) --
	(259.73, 73.36) --
	(258.38, 73.51) --
	(257.04, 73.68) --
	(255.69, 73.89) --
	(254.34, 74.12) --
	(252.99, 74.38) --
	(251.64, 74.67) --
	(250.30, 74.98) --
	(248.95, 75.32) --
	(247.60, 75.69) --
	(246.25, 76.08) --
	(244.90, 76.50) --
	(243.55, 76.94) --
	(242.21, 77.41) --
	(240.86, 77.91) --
	(239.51, 78.43) --
	(238.16, 78.97) --
	(236.81, 79.53) --
	(235.47, 80.12) --
	(234.12, 80.73) --
	(232.77, 81.37) --
	(231.42, 82.02) --
	(230.07, 82.70) --
	(228.73, 83.40) --
	(227.38, 84.12) --
	(226.03, 84.86) --
	(224.68, 85.61) --
	(223.33, 86.39) --
	(221.99, 87.19) --
	(220.64, 88.00) --
	(219.29, 88.83) --
	(217.94, 89.68) --
	(216.59, 90.55) --
	(215.24, 91.43) --
	(213.90, 92.32) --
	(212.55, 93.23) --
	(211.20, 94.16) --
	(209.85, 95.10) --
	(208.50, 96.05) --
	(207.16, 97.02) --
	(205.81, 97.99) --
	(204.46, 98.98) --
	(203.11, 99.98) --
	(201.76,100.99) --
	(200.42,102.02) --
	(199.07,103.05) --
	(197.72,104.08) --
	(196.37,105.13) --
	(195.02,106.18) --
	(193.68,107.25) --
	(192.33,108.31) --
	(190.98,109.39) --
	(189.63,110.46) --
	(188.28,111.55) --
	(186.94,112.63) --
	(185.59,113.72) --
	(184.24,114.81) --
	(182.89,115.91) --
	(181.54,117.01) --
	(180.19,118.10) --
	(178.85,119.20) --
	(177.50,120.30) --
	(176.15,121.39) --
	(174.80,122.49) --
	(173.45,123.58) --
	(172.11,124.67) --
	(170.76,125.76) --
	(169.41,126.84) --
	(168.06,127.92) --
	(166.71,128.99) --
	(165.37,130.06) --
	(164.02,131.12) --
	(162.67,132.18) --
	(161.32,133.23) --
	(159.97,134.27) --
	(158.63,135.30) --
	(157.28,136.32) --
	(155.93,137.34) --
	(154.58,138.34) --
	(153.23,139.34) --
	(151.89,140.32) --
	(150.54,141.29) --
	(149.19,142.25) --
	(147.84,143.20) --
	(146.49,144.13) --
	(145.14,145.05) --
	(143.80,145.96) --
	(142.45,146.85) --
	(141.10,147.72) --
	(139.75,148.59) --
	(138.40,149.43) --
	(137.06,150.26) --
	(135.71,151.08) --
	(134.36,151.87) --
	(133.01,152.65) --
	(131.66,153.41) --
	(130.32,154.15) --
	(128.97,154.88) --
	(127.62,155.58) --
	(126.27,156.27) --
	(124.92,156.93) --
	(123.58,157.58) --
	(122.23,158.20) --
	(120.88,158.80) --
	(119.53,159.39) --
	(118.18,159.95) --
	(116.83,160.49) --
	(115.49,161.00) --
	(114.14,161.50) --
	(112.79,161.97) --
	(111.44,162.42) --
	(110.09,162.84) --
	(108.75,163.25) --
	(107.40,163.63) --
	(106.05,163.98) --
	(104.70,164.31) --
	(103.35,164.62) --
	(102.01,164.90) --
	(100.66,165.15) --
	( 99.31,165.39) --
	( 97.96,165.59) --
	( 96.61,165.77) --
	( 95.27,165.93) --
	( 93.92,166.05) --
	( 92.57,166.16) --
	( 91.22,166.23) --
	( 89.87,166.28) --
	( 88.53,166.30) --
	( 87.18,166.30) --
	( 85.83,166.27) --
	( 84.48,166.21) --
	( 83.13,166.12) --
	( 81.78,166.01) --
	( 80.44,165.86) --
	( 79.09,165.69) --
	( 77.74,165.49) --
	( 76.39,165.26) --
	( 75.04,165.00) --
	( 73.70,164.71) --
	( 72.35,164.39) --
	( 71.00,164.04) --
	( 69.65,163.65) --
	( 68.30,163.23) --
	( 66.96,162.78) --
	( 65.61,162.29) --
	( 64.26,161.77) --
	( 62.91,161.20) --
	( 61.56,160.60) --
	( 60.22,159.95) --
	( 58.87,159.26) --
	( 57.52,158.51) --
	( 56.17,157.71) --
	( 54.82,156.85) --
	( 53.47,155.92) --
	( 52.13,154.90) --
	( 50.78,153.78) --
	( 49.43,152.53) --
	( 48.08,151.09) --
	( 46.73,149.38) --
	( 45.39,147.15) --
	( 44.04,141.73) --
	cycle;

\path[] ( 44.04,175.57) --
	( 45.39,175.57) --
	( 46.73,175.57) --
	( 48.08,175.57) --
	( 49.43,175.57) --
	( 50.78,175.57) --
	( 52.13,175.57) --
	( 53.47,175.57) --
	( 54.82,175.57) --
	( 56.17,175.57) --
	( 57.52,175.57) --
	( 58.87,175.57) --
	( 60.22,175.57) --
	( 61.56,175.57) --
	( 62.91,175.57) --
	( 64.26,175.57) --
	( 65.61,175.57) --
	( 66.96,175.57) --
	( 68.30,175.57) --
	( 69.65,175.57) --
	( 71.00,175.57) --
	( 72.35,175.57) --
	( 73.70,175.57) --
	( 75.04,175.57) --
	( 76.39,175.57) --
	( 77.74,175.57) --
	( 79.09,175.57) --
	( 80.44,175.57) --
	( 81.78,175.57) --
	( 83.13,175.57) --
	( 84.48,175.57) --
	( 85.83,175.57) --
	( 87.18,175.57) --
	( 88.53,175.57) --
	( 89.87,175.57) --
	( 91.22,175.57) --
	( 92.57,175.57) --
	( 93.92,175.57) --
	( 95.27,175.57) --
	( 96.61,175.57) --
	( 97.96,175.57) --
	( 99.31,175.57) --
	(100.66,175.57) --
	(102.01,175.57) --
	(103.35,175.57) --
	(104.70,175.57) --
	(106.05,175.57) --
	(107.40,175.57) --
	(108.75,175.57) --
	(110.09,175.57) --
	(111.44,175.57) --
	(112.79,175.57) --
	(114.14,175.57) --
	(115.49,175.57) --
	(116.83,175.57) --
	(118.18,175.57) --
	(119.53,175.57) --
	(120.88,175.57) --
	(122.23,175.57) --
	(123.58,175.57) --
	(124.92,175.57) --
	(126.27,175.57) --
	(127.62,175.57) --
	(128.97,175.57) --
	(130.32,175.57) --
	(131.66,175.57) --
	(133.01,175.57) --
	(134.36,175.57) --
	(135.71,175.57) --
	(137.06,175.57) --
	(138.40,175.57) --
	(139.75,175.57) --
	(141.10,175.57) --
	(142.45,175.57) --
	(143.80,175.57) --
	(145.14,175.57) --
	(146.49,175.57) --
	(147.84,175.57) --
	(149.19,175.57) --
	(150.54,175.57) --
	(151.89,175.57) --
	(153.23,175.57) --
	(154.58,175.57) --
	(155.93,175.57) --
	(157.28,175.57) --
	(158.63,175.57) --
	(159.97,175.57) --
	(161.32,175.57) --
	(162.67,175.57) --
	(164.02,175.57) --
	(165.37,175.57) --
	(166.71,175.57) --
	(168.06,175.57) --
	(169.41,175.57) --
	(170.76,175.57) --
	(172.11,175.57) --
	(173.45,175.57) --
	(174.80,175.57) --
	(176.15,175.57) --
	(177.50,175.57) --
	(178.85,175.57) --
	(180.19,175.57) --
	(181.54,175.57) --
	(182.89,175.57) --
	(184.24,175.57) --
	(185.59,175.57) --
	(186.94,175.57) --
	(188.28,175.57) --
	(189.63,175.57) --
	(190.98,175.57) --
	(192.33,175.57) --
	(193.68,175.57) --
	(195.02,175.57) --
	(196.37,175.57) --
	(197.72,175.57) --
	(199.07,175.57) --
	(200.42,175.57) --
	(201.76,175.57) --
	(203.11,175.57) --
	(204.46,175.57) --
	(205.81,175.57) --
	(207.16,175.57) --
	(208.50,175.57) --
	(209.85,175.57) --
	(211.20,175.57) --
	(212.55,175.57) --
	(213.90,175.57) --
	(215.24,175.57) --
	(216.59,175.57) --
	(217.94,175.57) --
	(219.29,175.57) --
	(220.64,175.57) --
	(221.99,175.57) --
	(223.33,175.57) --
	(224.68,175.57) --
	(226.03,175.57) --
	(227.38,175.57) --
	(228.73,175.57) --
	(230.07,175.57) --
	(231.42,175.57) --
	(232.77,175.57) --
	(234.12,175.57) --
	(235.47,175.57) --
	(236.81,175.57) --
	(238.16,175.57) --
	(239.51,175.57) --
	(240.86,175.57) --
	(242.21,175.57) --
	(243.55,175.57) --
	(244.90,175.57) --
	(246.25,175.57) --
	(247.60,175.57) --
	(248.95,175.57) --
	(250.30,175.57) --
	(251.64,175.57) --
	(252.99,175.57) --
	(254.34,175.57) --
	(255.69,175.57) --
	(257.04,175.57) --
	(258.38,175.57) --
	(259.73,175.57) --
	(261.08,175.57) --
	(262.43,175.57) --
	(263.78,175.57) --
	(265.12,175.57) --
	(266.47,175.57) --
	(267.82,175.57) --
	(269.17,175.57) --
	(270.52,175.57) --
	(271.86,175.57) --
	(273.21,175.57) --
	(274.56,175.57) --
	(275.91,175.57) --
	(277.26,175.57) --
	(278.60,175.57) --
	(279.95,175.57) --
	(281.30,175.57) --
	(282.65,175.57) --
	(284.00,175.57) --
	(285.35,175.57) --
	(286.69,175.57) --
	(288.04,175.57) --
	(289.39,175.57) --
	(290.74,175.57) --
	(292.09,175.57) --
	(293.43,175.57) --
	(294.78,175.57) --
	(296.13,175.57) --
	(297.48,175.57) --
	(298.83,175.57) --
	(300.17,175.57) --
	(301.52,175.57) --
	(302.87,175.57) --
	(304.22,175.57) --
	(305.57,175.57) --
	(306.91,175.57) --
	(308.26,175.57) --
	(309.61,175.57) --
	(310.96,175.57) --
	(312.31,175.57) --
	(313.65,175.57);

\path[] (313.65, 91.68) --
	(312.31, 90.66) --
	(310.96, 89.67) --
	(309.61, 88.70) --
	(308.26, 87.76) --
	(306.91, 86.84) --
	(305.57, 85.96) --
	(304.22, 85.10) --
	(302.87, 84.27) --
	(301.52, 83.47) --
	(300.17, 82.70) --
	(298.83, 81.96) --
	(297.48, 81.24) --
	(296.13, 80.56) --
	(294.78, 79.90) --
	(293.43, 79.27) --
	(292.09, 78.68) --
	(290.74, 78.11) --
	(289.39, 77.57) --
	(288.04, 77.07) --
	(286.69, 76.59) --
	(285.35, 76.14) --
	(284.00, 75.72) --
	(282.65, 75.34) --
	(281.30, 74.98) --
	(279.95, 74.66) --
	(278.60, 74.36) --
	(277.26, 74.09) --
	(275.91, 73.86) --
	(274.56, 73.65) --
	(273.21, 73.48) --
	(271.86, 73.33) --
	(270.52, 73.22) --
	(269.17, 73.13) --
	(267.82, 73.08) --
	(266.47, 73.05) --
	(265.12, 73.06) --
	(263.78, 73.09) --
	(262.43, 73.15) --
	(261.08, 73.24) --
	(259.73, 73.36) --
	(258.38, 73.51) --
	(257.04, 73.68) --
	(255.69, 73.89) --
	(254.34, 74.12) --
	(252.99, 74.38) --
	(251.64, 74.67) --
	(250.30, 74.98) --
	(248.95, 75.32) --
	(247.60, 75.69) --
	(246.25, 76.08) --
	(244.90, 76.50) --
	(243.55, 76.94) --
	(242.21, 77.41) --
	(240.86, 77.91) --
	(239.51, 78.43) --
	(238.16, 78.97) --
	(236.81, 79.53) --
	(235.47, 80.12) --
	(234.12, 80.73) --
	(232.77, 81.37) --
	(231.42, 82.02) --
	(230.07, 82.70) --
	(228.73, 83.40) --
	(227.38, 84.12) --
	(226.03, 84.86) --
	(224.68, 85.61) --
	(223.33, 86.39) --
	(221.99, 87.19) --
	(220.64, 88.00) --
	(219.29, 88.83) --
	(217.94, 89.68) --
	(216.59, 90.55) --
	(215.24, 91.43) --
	(213.90, 92.32) --
	(212.55, 93.23) --
	(211.20, 94.16) --
	(209.85, 95.10) --
	(208.50, 96.05) --
	(207.16, 97.02) --
	(205.81, 97.99) --
	(204.46, 98.98) --
	(203.11, 99.98) --
	(201.76,100.99) --
	(200.42,102.02) --
	(199.07,103.05) --
	(197.72,104.08) --
	(196.37,105.13) --
	(195.02,106.18) --
	(193.68,107.25) --
	(192.33,108.31) --
	(190.98,109.39) --
	(189.63,110.46) --
	(188.28,111.55) --
	(186.94,112.63) --
	(185.59,113.72) --
	(184.24,114.81) --
	(182.89,115.91) --
	(181.54,117.01) --
	(180.19,118.10) --
	(178.85,119.20) --
	(177.50,120.30) --
	(176.15,121.39) --
	(174.80,122.49) --
	(173.45,123.58) --
	(172.11,124.67) --
	(170.76,125.76) --
	(169.41,126.84) --
	(168.06,127.92) --
	(166.71,128.99) --
	(165.37,130.06) --
	(164.02,131.12) --
	(162.67,132.18) --
	(161.32,133.23) --
	(159.97,134.27) --
	(158.63,135.30) --
	(157.28,136.32) --
	(155.93,137.34) --
	(154.58,138.34) --
	(153.23,139.34) --
	(151.89,140.32) --
	(150.54,141.29) --
	(149.19,142.25) --
	(147.84,143.20) --
	(146.49,144.13) --
	(145.14,145.05) --
	(143.80,145.96) --
	(142.45,146.85) --
	(141.10,147.72) --
	(139.75,148.59) --
	(138.40,149.43) --
	(137.06,150.26) --
	(135.71,151.08) --
	(134.36,151.87) --
	(133.01,152.65) --
	(131.66,153.41) --
	(130.32,154.15) --
	(128.97,154.88) --
	(127.62,155.58) --
	(126.27,156.27) --
	(124.92,156.93) --
	(123.58,157.58) --
	(122.23,158.20) --
	(120.88,158.80) --
	(119.53,159.39) --
	(118.18,159.95) --
	(116.83,160.49) --
	(115.49,161.00) --
	(114.14,161.50) --
	(112.79,161.97) --
	(111.44,162.42) --
	(110.09,162.84) --
	(108.75,163.25) --
	(107.40,163.63) --
	(106.05,163.98) --
	(104.70,164.31) --
	(103.35,164.62) --
	(102.01,164.90) --
	(100.66,165.15) --
	( 99.31,165.39) --
	( 97.96,165.59) --
	( 96.61,165.77) --
	( 95.27,165.93) --
	( 93.92,166.05) --
	( 92.57,166.16) --
	( 91.22,166.23) --
	( 89.87,166.28) --
	( 88.53,166.30) --
	( 87.18,166.30) --
	( 85.83,166.27) --
	( 84.48,166.21) --
	( 83.13,166.12) --
	( 81.78,166.01) --
	( 80.44,165.86) --
	( 79.09,165.69) --
	( 77.74,165.49) --
	( 76.39,165.26) --
	( 75.04,165.00) --
	( 73.70,164.71) --
	( 72.35,164.39) --
	( 71.00,164.04) --
	( 69.65,163.65) --
	( 68.30,163.23) --
	( 66.96,162.78) --
	( 65.61,162.29) --
	( 64.26,161.77) --
	( 62.91,161.20) --
	( 61.56,160.60) --
	( 60.22,159.95) --
	( 58.87,159.26) --
	( 57.52,158.51) --
	( 56.17,157.71) --
	( 54.82,156.85) --
	( 53.47,155.92) --
	( 52.13,154.90) --
	( 50.78,153.78) --
	( 49.43,152.53) --
	( 48.08,151.09) --
	( 46.73,149.38) --
	( 45.39,147.15) --
	( 44.04,141.73);
\definecolor{fillColor}{RGB}{0,0,255}

\path[fill=fillColor,fill opacity=0.10] ( 44.04,141.73) --
	( 45.39,147.15) --
	( 46.73,149.38) --
	( 48.08,151.09) --
	( 49.43,152.53) --
	( 50.78,153.78) --
	( 52.13,154.90) --
	( 53.47,155.92) --
	( 54.82,156.85) --
	( 56.17,157.71) --
	( 57.52,158.51) --
	( 58.87,159.26) --
	( 60.22,159.95) --
	( 61.56,160.60) --
	( 62.91,161.20) --
	( 64.26,161.77) --
	( 65.61,162.29) --
	( 66.96,162.78) --
	( 68.30,163.23) --
	( 69.65,163.65) --
	( 71.00,164.04) --
	( 72.35,164.39) --
	( 73.70,164.71) --
	( 75.04,165.00) --
	( 76.39,165.26) --
	( 77.74,165.49) --
	( 79.09,165.69) --
	( 80.44,165.86) --
	( 81.78,166.01) --
	( 83.13,166.12) --
	( 84.48,166.21) --
	( 85.83,166.27) --
	( 87.18,166.30) --
	( 88.53,166.30) --
	( 89.87,166.28) --
	( 91.22,166.23) --
	( 92.57,166.16) --
	( 93.92,166.05) --
	( 95.27,165.93) --
	( 96.61,165.77) --
	( 97.96,165.59) --
	( 99.31,165.39) --
	(100.66,165.15) --
	(102.01,164.90) --
	(103.35,164.62) --
	(104.70,164.31) --
	(106.05,163.98) --
	(107.40,163.63) --
	(108.75,163.25) --
	(110.09,162.84) --
	(111.44,162.42) --
	(112.79,161.97) --
	(114.14,161.50) --
	(115.49,161.00) --
	(116.83,160.49) --
	(118.18,159.95) --
	(119.53,159.39) --
	(120.88,158.80) --
	(122.23,158.20) --
	(123.58,157.58) --
	(124.92,156.93) --
	(126.27,156.27) --
	(127.62,155.58) --
	(128.97,154.88) --
	(130.32,154.15) --
	(131.66,153.41) --
	(133.01,152.65) --
	(134.36,151.87) --
	(135.71,151.08) --
	(137.06,150.26) --
	(138.40,149.43) --
	(139.75,148.59) --
	(141.10,147.72) --
	(142.45,146.85) --
	(143.80,145.96) --
	(145.14,145.05) --
	(146.49,144.13) --
	(147.84,143.20) --
	(149.19,142.25) --
	(150.54,141.29) --
	(151.89,140.32) --
	(153.23,139.34) --
	(154.58,138.34) --
	(155.93,137.34) --
	(157.28,136.32) --
	(158.63,135.30) --
	(159.97,134.27) --
	(161.32,133.23) --
	(162.67,132.18) --
	(164.02,131.12) --
	(165.37,130.06) --
	(166.71,128.99) --
	(168.06,127.92) --
	(169.41,126.84) --
	(170.76,125.76) --
	(172.11,124.67) --
	(173.45,123.58) --
	(174.80,122.49) --
	(176.15,121.39) --
	(177.50,120.30) --
	(178.85,119.20) --
	(180.19,118.10) --
	(181.54,117.01) --
	(182.89,115.91) --
	(184.24,114.81) --
	(185.59,113.72) --
	(186.94,112.63) --
	(188.28,111.55) --
	(189.63,110.46) --
	(190.98,109.39) --
	(192.33,108.31) --
	(193.68,107.25) --
	(195.02,106.18) --
	(196.37,105.13) --
	(197.72,104.08) --
	(199.07,103.05) --
	(200.42,102.02) --
	(201.76,100.99) --
	(203.11, 99.98) --
	(204.46, 98.98) --
	(205.81, 97.99) --
	(207.16, 97.02) --
	(208.50, 96.05) --
	(209.85, 95.10) --
	(211.20, 94.16) --
	(212.55, 93.23) --
	(213.90, 92.32) --
	(215.24, 91.43) --
	(216.59, 90.55) --
	(217.94, 89.68) --
	(219.29, 88.83) --
	(220.64, 88.00) --
	(221.99, 87.19) --
	(223.33, 86.39) --
	(224.68, 85.61) --
	(226.03, 84.86) --
	(227.38, 84.12) --
	(228.73, 83.40) --
	(230.07, 82.70) --
	(231.42, 82.02) --
	(232.77, 81.37) --
	(234.12, 80.73) --
	(235.47, 80.12) --
	(236.81, 79.53) --
	(238.16, 78.97) --
	(239.51, 78.43) --
	(240.86, 77.91) --
	(242.21, 77.41) --
	(243.55, 76.94) --
	(244.90, 76.50) --
	(246.25, 76.08) --
	(247.60, 75.69) --
	(248.95, 75.32) --
	(250.30, 74.98) --
	(251.64, 74.67) --
	(252.99, 74.38) --
	(254.34, 74.12) --
	(255.69, 73.89) --
	(257.04, 73.68) --
	(258.38, 73.51) --
	(259.73, 73.36) --
	(261.08, 73.24) --
	(262.43, 73.15) --
	(263.78, 73.09) --
	(265.12, 73.06) --
	(266.47, 73.05) --
	(267.82, 73.08) --
	(269.17, 73.13) --
	(270.52, 73.22) --
	(271.86, 73.33) --
	(273.21, 73.48) --
	(274.56, 73.65) --
	(275.91, 73.86) --
	(277.26, 74.09) --
	(278.60, 74.36) --
	(279.95, 74.66) --
	(281.30, 74.98) --
	(282.65, 75.34) --
	(284.00, 75.72) --
	(285.35, 76.14) --
	(286.69, 76.59) --
	(288.04, 77.07) --
	(289.39, 77.57) --
	(290.74, 78.11) --
	(292.09, 78.68) --
	(293.43, 79.27) --
	(294.78, 79.90) --
	(296.13, 80.56) --
	(297.48, 81.24) --
	(298.83, 81.96) --
	(300.17, 82.70) --
	(301.52, 83.47) --
	(302.87, 84.27) --
	(304.22, 85.10) --
	(305.57, 85.96) --
	(306.91, 86.84) --
	(308.26, 87.76) --
	(309.61, 88.70) --
	(310.96, 89.67) --
	(312.31, 90.66) --
	(313.65, 91.68) --
	(313.65, 40.21) --
	(312.31, 40.21) --
	(310.96, 40.21) --
	(309.61, 40.21) --
	(308.26, 40.21) --
	(306.91, 40.21) --
	(305.57, 40.21) --
	(304.22, 40.21) --
	(302.87, 40.21) --
	(301.52, 40.21) --
	(300.17, 40.21) --
	(298.83, 40.21) --
	(297.48, 40.21) --
	(296.13, 40.21) --
	(294.78, 40.21) --
	(293.43, 40.21) --
	(292.09, 40.21) --
	(290.74, 40.21) --
	(289.39, 40.21) --
	(288.04, 40.21) --
	(286.69, 40.21) --
	(285.35, 40.21) --
	(284.00, 40.21) --
	(282.65, 40.21) --
	(281.30, 40.21) --
	(279.95, 40.21) --
	(278.60, 40.21) --
	(277.26, 40.21) --
	(275.91, 40.21) --
	(274.56, 40.21) --
	(273.21, 40.21) --
	(271.86, 40.21) --
	(270.52, 40.21) --
	(269.17, 40.21) --
	(267.82, 40.21) --
	(266.47, 40.21) --
	(265.12, 40.21) --
	(263.78, 40.21) --
	(262.43, 40.21) --
	(261.08, 40.21) --
	(259.73, 40.21) --
	(258.38, 40.21) --
	(257.04, 40.21) --
	(255.69, 40.21) --
	(254.34, 40.21) --
	(252.99, 40.21) --
	(251.64, 40.21) --
	(250.30, 40.21) --
	(248.95, 40.21) --
	(247.60, 40.21) --
	(246.25, 40.21) --
	(244.90, 40.21) --
	(243.55, 40.21) --
	(242.21, 40.21) --
	(240.86, 40.21) --
	(239.51, 40.21) --
	(238.16, 40.21) --
	(236.81, 40.21) --
	(235.47, 40.21) --
	(234.12, 40.21) --
	(232.77, 40.21) --
	(231.42, 40.21) --
	(230.07, 40.21) --
	(228.73, 40.21) --
	(227.38, 40.21) --
	(226.03, 40.21) --
	(224.68, 40.21) --
	(223.33, 40.21) --
	(221.99, 40.21) --
	(220.64, 40.21) --
	(219.29, 40.21) --
	(217.94, 40.21) --
	(216.59, 40.21) --
	(215.24, 40.21) --
	(213.90, 40.21) --
	(212.55, 40.21) --
	(211.20, 40.21) --
	(209.85, 40.21) --
	(208.50, 40.21) --
	(207.16, 40.21) --
	(205.81, 40.21) --
	(204.46, 40.21) --
	(203.11, 40.21) --
	(201.76, 40.21) --
	(200.42, 40.21) --
	(199.07, 40.21) --
	(197.72, 40.21) --
	(196.37, 40.21) --
	(195.02, 40.21) --
	(193.68, 40.21) --
	(192.33, 40.21) --
	(190.98, 40.21) --
	(189.63, 40.21) --
	(188.28, 40.21) --
	(186.94, 40.21) --
	(185.59, 40.21) --
	(184.24, 40.21) --
	(182.89, 40.21) --
	(181.54, 40.21) --
	(180.19, 40.21) --
	(178.85, 40.21) --
	(177.50, 40.21) --
	(176.15, 40.21) --
	(174.80, 40.21) --
	(173.45, 40.21) --
	(172.11, 40.21) --
	(170.76, 40.21) --
	(169.41, 40.21) --
	(168.06, 40.21) --
	(166.71, 40.21) --
	(165.37, 40.21) --
	(164.02, 40.21) --
	(162.67, 40.21) --
	(161.32, 40.21) --
	(159.97, 40.21) --
	(158.63, 40.21) --
	(157.28, 40.21) --
	(155.93, 40.21) --
	(154.58, 40.21) --
	(153.23, 40.21) --
	(151.89, 40.21) --
	(150.54, 40.21) --
	(149.19, 40.21) --
	(147.84, 40.21) --
	(146.49, 40.21) --
	(145.14, 40.21) --
	(143.80, 40.21) --
	(142.45, 40.21) --
	(141.10, 40.21) --
	(139.75, 40.21) --
	(138.40, 40.21) --
	(137.06, 40.21) --
	(135.71, 40.21) --
	(134.36, 40.21) --
	(133.01, 40.21) --
	(131.66, 40.21) --
	(130.32, 40.21) --
	(128.97, 40.21) --
	(127.62, 40.21) --
	(126.27, 40.21) --
	(124.92, 40.21) --
	(123.58, 40.21) --
	(122.23, 40.21) --
	(120.88, 40.21) --
	(119.53, 40.21) --
	(118.18, 40.21) --
	(116.83, 40.21) --
	(115.49, 40.21) --
	(114.14, 40.21) --
	(112.79, 40.21) --
	(111.44, 40.21) --
	(110.09, 40.21) --
	(108.75, 40.21) --
	(107.40, 40.21) --
	(106.05, 40.21) --
	(104.70, 40.21) --
	(103.35, 40.21) --
	(102.01, 40.21) --
	(100.66, 40.21) --
	( 99.31, 40.21) --
	( 97.96, 40.21) --
	( 96.61, 40.21) --
	( 95.27, 40.21) --
	( 93.92, 40.21) --
	( 92.57, 40.21) --
	( 91.22, 40.21) --
	( 89.87, 40.21) --
	( 88.53, 40.21) --
	( 87.18, 40.21) --
	( 85.83, 40.21) --
	( 84.48, 40.21) --
	( 83.13, 40.21) --
	( 81.78, 40.21) --
	( 80.44, 40.21) --
	( 79.09, 40.21) --
	( 77.74, 40.21) --
	( 76.39, 40.21) --
	( 75.04, 40.21) --
	( 73.70, 40.21) --
	( 72.35, 40.21) --
	( 71.00, 40.21) --
	( 69.65, 40.21) --
	( 68.30, 40.21) --
	( 66.96, 40.21) --
	( 65.61, 40.21) --
	( 64.26, 40.21) --
	( 62.91, 40.21) --
	( 61.56, 40.21) --
	( 60.22, 40.21) --
	( 58.87, 40.21) --
	( 57.52, 40.21) --
	( 56.17, 40.21) --
	( 54.82, 40.21) --
	( 53.47, 40.21) --
	( 52.13, 40.21) --
	( 50.78, 40.21) --
	( 49.43, 40.21) --
	( 48.08, 40.21) --
	( 46.73, 40.21) --
	( 45.39, 40.21) --
	( 44.04, 40.21) --
	cycle;

\path[] ( 44.04,141.73) --
	( 45.39,147.15) --
	( 46.73,149.38) --
	( 48.08,151.09) --
	( 49.43,152.53) --
	( 50.78,153.78) --
	( 52.13,154.90) --
	( 53.47,155.92) --
	( 54.82,156.85) --
	( 56.17,157.71) --
	( 57.52,158.51) --
	( 58.87,159.26) --
	( 60.22,159.95) --
	( 61.56,160.60) --
	( 62.91,161.20) --
	( 64.26,161.77) --
	( 65.61,162.29) --
	( 66.96,162.78) --
	( 68.30,163.23) --
	( 69.65,163.65) --
	( 71.00,164.04) --
	( 72.35,164.39) --
	( 73.70,164.71) --
	( 75.04,165.00) --
	( 76.39,165.26) --
	( 77.74,165.49) --
	( 79.09,165.69) --
	( 80.44,165.86) --
	( 81.78,166.01) --
	( 83.13,166.12) --
	( 84.48,166.21) --
	( 85.83,166.27) --
	( 87.18,166.30) --
	( 88.53,166.30) --
	( 89.87,166.28) --
	( 91.22,166.23) --
	( 92.57,166.16) --
	( 93.92,166.05) --
	( 95.27,165.93) --
	( 96.61,165.77) --
	( 97.96,165.59) --
	( 99.31,165.39) --
	(100.66,165.15) --
	(102.01,164.90) --
	(103.35,164.62) --
	(104.70,164.31) --
	(106.05,163.98) --
	(107.40,163.63) --
	(108.75,163.25) --
	(110.09,162.84) --
	(111.44,162.42) --
	(112.79,161.97) --
	(114.14,161.50) --
	(115.49,161.00) --
	(116.83,160.49) --
	(118.18,159.95) --
	(119.53,159.39) --
	(120.88,158.80) --
	(122.23,158.20) --
	(123.58,157.58) --
	(124.92,156.93) --
	(126.27,156.27) --
	(127.62,155.58) --
	(128.97,154.88) --
	(130.32,154.15) --
	(131.66,153.41) --
	(133.01,152.65) --
	(134.36,151.87) --
	(135.71,151.08) --
	(137.06,150.26) --
	(138.40,149.43) --
	(139.75,148.59) --
	(141.10,147.72) --
	(142.45,146.85) --
	(143.80,145.96) --
	(145.14,145.05) --
	(146.49,144.13) --
	(147.84,143.20) --
	(149.19,142.25) --
	(150.54,141.29) --
	(151.89,140.32) --
	(153.23,139.34) --
	(154.58,138.34) --
	(155.93,137.34) --
	(157.28,136.32) --
	(158.63,135.30) --
	(159.97,134.27) --
	(161.32,133.23) --
	(162.67,132.18) --
	(164.02,131.12) --
	(165.37,130.06) --
	(166.71,128.99) --
	(168.06,127.92) --
	(169.41,126.84) --
	(170.76,125.76) --
	(172.11,124.67) --
	(173.45,123.58) --
	(174.80,122.49) --
	(176.15,121.39) --
	(177.50,120.30) --
	(178.85,119.20) --
	(180.19,118.10) --
	(181.54,117.01) --
	(182.89,115.91) --
	(184.24,114.81) --
	(185.59,113.72) --
	(186.94,112.63) --
	(188.28,111.55) --
	(189.63,110.46) --
	(190.98,109.39) --
	(192.33,108.31) --
	(193.68,107.25) --
	(195.02,106.18) --
	(196.37,105.13) --
	(197.72,104.08) --
	(199.07,103.05) --
	(200.42,102.02) --
	(201.76,100.99) --
	(203.11, 99.98) --
	(204.46, 98.98) --
	(205.81, 97.99) --
	(207.16, 97.02) --
	(208.50, 96.05) --
	(209.85, 95.10) --
	(211.20, 94.16) --
	(212.55, 93.23) --
	(213.90, 92.32) --
	(215.24, 91.43) --
	(216.59, 90.55) --
	(217.94, 89.68) --
	(219.29, 88.83) --
	(220.64, 88.00) --
	(221.99, 87.19) --
	(223.33, 86.39) --
	(224.68, 85.61) --
	(226.03, 84.86) --
	(227.38, 84.12) --
	(228.73, 83.40) --
	(230.07, 82.70) --
	(231.42, 82.02) --
	(232.77, 81.37) --
	(234.12, 80.73) --
	(235.47, 80.12) --
	(236.81, 79.53) --
	(238.16, 78.97) --
	(239.51, 78.43) --
	(240.86, 77.91) --
	(242.21, 77.41) --
	(243.55, 76.94) --
	(244.90, 76.50) --
	(246.25, 76.08) --
	(247.60, 75.69) --
	(248.95, 75.32) --
	(250.30, 74.98) --
	(251.64, 74.67) --
	(252.99, 74.38) --
	(254.34, 74.12) --
	(255.69, 73.89) --
	(257.04, 73.68) --
	(258.38, 73.51) --
	(259.73, 73.36) --
	(261.08, 73.24) --
	(262.43, 73.15) --
	(263.78, 73.09) --
	(265.12, 73.06) --
	(266.47, 73.05) --
	(267.82, 73.08) --
	(269.17, 73.13) --
	(270.52, 73.22) --
	(271.86, 73.33) --
	(273.21, 73.48) --
	(274.56, 73.65) --
	(275.91, 73.86) --
	(277.26, 74.09) --
	(278.60, 74.36) --
	(279.95, 74.66) --
	(281.30, 74.98) --
	(282.65, 75.34) --
	(284.00, 75.72) --
	(285.35, 76.14) --
	(286.69, 76.59) --
	(288.04, 77.07) --
	(289.39, 77.57) --
	(290.74, 78.11) --
	(292.09, 78.68) --
	(293.43, 79.27) --
	(294.78, 79.90) --
	(296.13, 80.56) --
	(297.48, 81.24) --
	(298.83, 81.96) --
	(300.17, 82.70) --
	(301.52, 83.47) --
	(302.87, 84.27) --
	(304.22, 85.10) --
	(305.57, 85.96) --
	(306.91, 86.84) --
	(308.26, 87.76) --
	(309.61, 88.70) --
	(310.96, 89.67) --
	(312.31, 90.66) --
	(313.65, 91.68);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (313.65, 29.64) {$T$};
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 35.95,171.77) {$\bar \Lambda$};
\definecolor{drawColor}{RGB}{0,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 35.95,137.93) {$\Lambda$};
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] ( 67.44, 95.74) -- ( 74.56,102.87);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] ( 67.44,102.87) -- ( 74.56, 95.74);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (113.06,117.88) -- (120.19,125.01);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (113.06,125.01) -- (120.19,117.88);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (188.76,159.55) -- (195.89,166.68);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (188.76,166.68) -- (195.89,159.55);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (242.85, 64.90) -- (249.98, 72.02);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (242.85, 72.02) -- (249.98, 64.90);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (280.23, 85.92) -- (287.36, 93.05);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (280.23, 93.05) -- (287.36, 85.92);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] ( 71.00, 99.31) circle (  3.56);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (116.62,121.45) circle (  3.56);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (246.41, 68.46) circle (  3.56);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.3pt,dash pattern=on 1pt off 3pt ,line join=round] ( 71.00, 99.31) -- ( 71.00, 42.92);

\path[draw=drawColor,line width= 0.3pt,dash pattern=on 1pt off 3pt ,line join=round] (116.62,121.45) -- (116.62, 42.92);

\path[draw=drawColor,line width= 0.3pt,dash pattern=on 1pt off 3pt ,line join=round] (192.33,163.11) -- (192.33, 42.92);

\path[draw=drawColor,line width= 0.3pt,dash pattern=on 1pt off 3pt ,line join=round] (246.41, 68.46) -- (246.41, 42.92);

\path[draw=drawColor,line width= 0.3pt,dash pattern=on 1pt off 3pt ,line join=round] (283.79, 89.48) -- (283.79, 42.92);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 71.00, 36.15) -- ( 71.00, 44.27);

\path[draw=drawColor,line width= 1.1pt,line join=round] (116.62, 36.15) -- (116.62, 44.27);

\path[draw=drawColor,line width= 1.1pt,line join=round] (246.41, 36.15) -- (246.41, 44.27);
\definecolor{drawColor}{RGB}{0,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 71.00, 25.58) {$T_{1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (116.62, 25.58) {$T_{2}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (246.41, 25.58) {$T_{3}$};
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 62.91, 44.53) {$\bar T_{1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (108.54, 44.53) {$\bar T_{2}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (184.24, 44.53) {$\bar T_{3}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (238.33, 44.53) {$\bar T_{4}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (275.70, 44.53) {$\bar T_{5}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (188.28,  7.64) {Time};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,116.01) {Marks};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (262.29,165.90) rectangle (355.85,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (271.46,189.52) -- (278.58,196.65);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (271.46,196.65) -- (278.58,189.52);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (271.46,189.52) -- (278.58,196.65);

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (271.46,196.65) -- (278.58,189.52);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (275.02,178.63) circle (  3.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round] (275.02,178.63) circle (  3.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (287.75,190.05) {Proposed Marks};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (287.75,175.60) {Accepted Marks};
\end{scope}
\end{tikzpicture}